---
title: "World Values Survey Analysis"
author: "Franchesca Garcia"
format: 
  pdf:
    warning: false
    message: false
    toc: true
    toc-depth: 2
editor: visual
---

```{r load in packages}
library(tidyverse)
library(car)
library(pROC)
library(caret)
library(AICcmodavg)
library(performance)
```

```{r load in data}
WVS <- read_csv("WVS.csv")

# cast predictors to factors
WVS$gender <- factor(WVS$gender)
WVS$religion <- factor(WVS$religion)
WVS$degree <- factor(WVS$degree)
WVS$country <- factor(WVS$country)
WVS$age <- factor(WVS$age)

# compute numbers of trials
WVS$total <- WVS$yes + WVS$no

# drop rows with 0 trials
WVS <- WVS |>
  filter(total > 0)

# compute proportions
WVS$prop <- WVS$yes/WVS$total
```

```{r ungroup data}
# ungroup data for roc-auc analysis
yes <- WVS |>
  uncount(yes) |>
  mutate(resp = 1)
no <- WVS |>
  uncount(no) |>
  mutate(resp = 0)

WVS_ungrouped <- bind_rows(yes, no) |>
  select(-c(yes, no))

rm(yes, no)
```

## Model 1: logit, all predictors

### 1.

```{r model1 fit}
model1 <- glm(cbind(yes, no) ~ gender + religion + degree + country + age, family = binomial(link="logit"), data = WVS)

summary(model1)
```

### 2.

```{r model1 gof test}
pchisq(deviance(model1), df.residual(model1), lower.tail = FALSE)
```

### 3.

```{r model1 sample size check}
model1_df = data_frame()

WVS$model1_preds <- fitted(model1, type="response")
model1_df = data_frame(yes = WVS$yes,
                       no = WVS$no,
                       pred_yes = WVS$total * WVS$model1_preds,
                       pred_no = WVS$total * (1 - WVS$model1_preds))

sum(model1_df$pred_yes < 5)
sum(model1_df$pred_yes < 1)
sum(model1_df$pred_no < 5)
sum(model1_df$pred_no < 1)
```

### 4.

```{r model1 multicolineaity check}
vif(model1)
```

### 5.

```{r model1 residuals check}
residualPlots(model1, type = "deviance", smooth=FALSE)
residualPlots(model1, type = "pearson", smooth=FALSE)
```

### 6.

```{r model1 influential obs check}
p <- length(coef(model1))
n <- nobs(model1)
leverage_cutoff <-  (3*p)/n
cooks_cutoff <- 2/sqrt(n)

model1_inf <- data.frame(
  obs = seq(n),
  studentized = rstudent(model1),
  hat = hatvalues(model1),
  cooks = cooks.distance(model1)
)

model1_inf |>
  ggplot(aes(x = obs, y = studentized)) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.25) +
  geom_hline(yintercept = -3, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_hline(yintercept = 3, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_segment(aes(xend = obs, y = 0, yend = studentized), linewidth = 0.25) +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("Studentized residuals") + 
  theme_bw()

model1_inf |>
  ggplot(aes(x = obs, y = hat)) +
  geom_segment(aes(xend = obs, y = 0, yend = hat), linewidth = 0.25) +
  geom_hline(yintercept = leverage_cutoff, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("hat-values") + 
  theme_bw()

model1_inf |>
  ggplot(aes(x = obs, y = cooks)) +
  geom_segment(aes(xend = obs, y = 0, yend = cooks), linewidth = 0.25) +
  geom_hline(yintercept = cooks_cutoff, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("Cook's distance") + 
  theme_bw()

model1_inf |>
  filter(abs(studentized) > 3 |
         hat > leverage_cutoff |
         cooks > cooks_cutoff
  )
```

### 7.

```{r model1 overdispersion check}
check_overdispersion(model1)
```

## Model 2: logit, stepwise selected predictors

### 8.

```{r model2 fit}
full_model <- glm(cbind(yes, no) ~ (gender + religion + degree + country + age)^2, family = binomial(link="logit"), data = WVS)

model2 <- step(full_model, direction = "both")
rm(full_model)

summary(model2)
```

### 9.

```{r model2 gof test}
pchisq(deviance(model2), df.residual(model2), lower.tail = FALSE)
```

### 10.

```{r model2 sample size check}
model2_df = data_frame()

WVS$model2_preds <- fitted(model2, type="response")
model2_df = data_frame(yes = WVS$yes,
                       no = WVS$no,
                       pred_yes = WVS$total * WVS$model2_preds,
                       pred_no = WVS$total * (1 - WVS$model2_preds))

sum(model2_df$pred_yes < 5)
sum(model2_df$pred_yes < 1)
sum(model2_df$pred_no < 5)
sum(model2_df$pred_no < 1)
```

### 11.

```{r model2 multicolineaity check}
vif(model2)
```

### 12.

```{r model2 residuals check}
residualPlots(model2, type = "deviance", smooth=FALSE)
residualPlots(model2, type = "pearson", smooth=FALSE)
```

### 13.

```{r model2 influential obs check}
p <- length(coef(model2))
n <- nobs(model2)
leverage_cutoff <-  (3*p)/n
cooks_cutoff <- 2/sqrt(n)

model2_inf <- data.frame(
  obs = seq(n),
  studentized = rstudent(model2),
  hat = hatvalues(model2),
  cooks = cooks.distance(model2)
)

model2_inf |>
  ggplot(aes(x = obs, y = studentized)) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.25) +
  geom_hline(yintercept = -3, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_hline(yintercept = 3, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_segment(aes(xend = obs, y = 0, yend = studentized), linewidth = 0.25) +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("Studentized residuals") + 
  theme_bw()

model2_inf |>
  ggplot(aes(x = obs, y = hat)) +
  geom_segment(aes(xend = obs, y = 0, yend = hat), linewidth = 0.25) +
  geom_hline(yintercept = leverage_cutoff, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("hat-values") + 
  theme_bw()

model2_inf |>
  ggplot(aes(x = obs, y = cooks)) +
  geom_segment(aes(xend = obs, y = 0, yend = cooks), linewidth = 0.25) +
  geom_hline(yintercept = cooks_cutoff, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("Cook's distance") + 
  theme_bw()

model2_inf |>
  filter(abs(studentized) > 3 |
         hat > leverage_cutoff |
         cooks > cooks_cutoff
  )
```

### 14.

```{r model2 overdispersion check}
check_overdispersion(model2)
```

## Model 3: probit, all predictors

### 15.

```{r model3 fit}
model3 <- glm(cbind(yes, no) ~ gender + religion + degree + country + age, family = binomial(link="probit"), data = WVS)

summary(model3)
```

### 16.

```{r model3 gof test}
pchisq(deviance(model3), df.residual(model3), lower.tail = FALSE)
```

### 17.

```{r model3 sample size check}
model3_df = data_frame()

WVS$model3_preds <- fitted(model3, type="response")
model3_df = data_frame(yes = WVS$yes,
                       no = WVS$no,
                       pred_yes = WVS$total * WVS$model3_preds,
                       pred_no = WVS$total * (1 - WVS$model3_preds))

sum(model3_df$pred_yes < 5)
sum(model3_df$pred_yes < 1)
sum(model3_df$pred_no < 5)
sum(model3_df$pred_no < 1)
```

### 18.

```{r model3 multicolineaity check}
vif(model3)
```

### 19.

```{r model3 residuals check}
residualPlots(model3, type = "deviance", smooth=FALSE)
residualPlots(model3, type = "pearson", smooth=FALSE)
```

### 20.

```{r model3 influential obs check}
p <- length(coef(model3))
n <- nobs(model3)
leverage_cutoff <-  (3*p)/n
cooks_cutoff <- 2/sqrt(n)

model3_inf <- data.frame(
  obs = seq(n),
  studentized = rstudent(model3),
  hat = hatvalues(model3),
  cooks = cooks.distance(model3)
)

model3_inf |>
  ggplot(aes(x = obs, y = studentized)) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.25) +
  geom_hline(yintercept = -3, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_hline(yintercept = 3, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_segment(aes(xend = obs, y = 0, yend = studentized), linewidth = 0.25) +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("Studentized residuals") + 
  theme_bw()

model3_inf |>
  ggplot(aes(x = obs, y = hat)) +
  geom_segment(aes(xend = obs, y = 0, yend = hat), linewidth = 0.25) +
  geom_hline(yintercept = leverage_cutoff, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("hat-values") + 
  theme_bw()

model3_inf |>
  ggplot(aes(x = obs, y = cooks)) +
  geom_segment(aes(xend = obs, y = 0, yend = cooks), linewidth = 0.25) +
  geom_hline(yintercept = cooks_cutoff, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("Cook's distance") + 
  theme_bw()

model3_inf |>
  filter(abs(studentized) > 3 |
         hat > leverage_cutoff |
         cooks > cooks_cutoff
  )
```

### 21.

```{r model3 overdispersion check}
check_overdispersion(model3)
```

## Model 4: probit, stepwise selected predictors

### 22.

```{r model4 fit}
full_model <- glm(cbind(yes, no) ~ (gender + religion + degree + country + age)^2, family = binomial(link="probit"), data = WVS)

model4 <- step(full_model, direction = "both")
rm(full_model)

summary(model4)
```

### 23.

```{r model4 gof test}
pchisq(deviance(model4), df.residual(model4), lower.tail = FALSE)
```

### 24.

```{r model4 sample size check}
model4_df = data_frame()

WVS$model4_preds <- fitted(model4, type="response")
model4_df = data_frame(yes = WVS$yes,
                       no = WVS$no,
                       pred_yes = WVS$total * WVS$model4_preds,
                       pred_no = WVS$total * (1 - WVS$model4_preds))

sum(model4_df$pred_yes < 5)
sum(model4_df$pred_yes < 1)
sum(model4_df$pred_no < 5)
sum(model4_df$pred_no < 1)
```

### 25.

```{r model4 multicolineaity check}
vif(model4)
```

### 26.

```{r model4 residuals check}
residualPlots(model4, type = "deviance", smooth=FALSE)
residualPlots(model4, type = "pearson", smooth=FALSE)
```

### 27.

```{r model4 influential obs check}
p <- length(coef(model4))
n <- nobs(model4)
leverage_cutoff <-  (3*p)/n
cooks_cutoff <- 2/sqrt(n)

model4_inf <- data.frame(
  obs = seq(n),
  studentized = rstudent(model4),
  hat = hatvalues(model4),
  cooks = cooks.distance(model4)
)

model4_inf |>
  ggplot(aes(x = obs, y = studentized)) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.25) +
  geom_hline(yintercept = -3, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_hline(yintercept = 3, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_segment(aes(xend = obs, y = 0, yend = studentized), linewidth = 0.25) +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("Studentized residuals") + 
  theme_bw()

model4_inf |>
  ggplot(aes(x = obs, y = hat)) +
  geom_segment(aes(xend = obs, y = 0, yend = hat), linewidth = 0.25) +
  geom_hline(yintercept = leverage_cutoff, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("hat-values") + 
  theme_bw()

model4_inf |>
  ggplot(aes(x = obs, y = cooks)) +
  geom_segment(aes(xend = obs, y = 0, yend = cooks), linewidth = 0.25) +
  geom_hline(yintercept = cooks_cutoff, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("Cook's distance") + 
  theme_bw()

model4_inf |>
  filter(abs(studentized) > 3 |
         hat > leverage_cutoff |
         cooks > cooks_cutoff
  )
```

### 28.

```{r model4 overdispersion check}
check_overdispersion(model4)
```

## Model 5: cloglog, all predictors

### 29.

```{r model5 fit}
model5 <- glm(cbind(yes, no) ~ gender + religion + degree + country + age, family = binomial(link="cloglog"), data = WVS)

summary(model5)
```

### 30.

```{r model5 gof test}
pchisq(deviance(model5), df.residual(model5), lower.tail = FALSE)
```

### 31.

```{r model5 sample size check}
model5_df = data_frame()

WVS$model5_preds <- fitted(model5, type="response")
model5_df = data_frame(yes = WVS$yes,
                       no = WVS$no,
                       pred_yes = WVS$total * WVS$model5_preds,
                       pred_no = WVS$total * (1 - WVS$model5_preds))

sum(model5_df$pred_yes < 5)
sum(model5_df$pred_yes < 1)
sum(model5_df$pred_no < 5)
sum(model5_df$pred_no < 1)
```

### 32.

```{r model5 multicolineaity check}
vif(model5)
```

### 33.

```{r model5 residuals check}
residualPlots(model5, type = "deviance", smooth=FALSE)
residualPlots(model5, type = "pearson", smooth=FALSE)
```

### 34.

```{r model5 influential obs check}
p <- length(coef(model5))
n <- nobs(model5)
leverage_cutoff <-  (3*p)/n
cooks_cutoff <- 2/sqrt(n)

model5_inf <- data.frame(
  obs = seq(n),
  studentized = rstudent(model5),
  hat = hatvalues(model5),
  cooks = cooks.distance(model5)
)

model5_inf |>
  ggplot(aes(x = obs, y = studentized)) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.25) +
  geom_hline(yintercept = -3, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_hline(yintercept = 3, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_segment(aes(xend = obs, y = 0, yend = studentized), linewidth = 0.25) +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("Studentized residuals") + 
  theme_bw()

model5_inf |>
  ggplot(aes(x = obs, y = hat)) +
  geom_segment(aes(xend = obs, y = 0, yend = hat), linewidth = 0.25) +
  geom_hline(yintercept = leverage_cutoff, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("hat-values") + 
  theme_bw()

model5_inf |>
  ggplot(aes(x = obs, y = cooks)) +
  geom_segment(aes(xend = obs, y = 0, yend = cooks), linewidth = 0.25) +
  geom_hline(yintercept = cooks_cutoff, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("Cook's distance") + 
  theme_bw()

model5_inf |>
  filter(abs(studentized) > 3 |
         hat > leverage_cutoff |
         cooks > cooks_cutoff
  )
```

### 35.

```{r model5 overdispersion check}
check_overdispersion(model5)
```

## Model 6: cloglog, stepwise selected predictors

### 36.

```{r model6 fit}
full_model <- glm(cbind(yes, no) ~ (gender + religion + degree + country + age)^2, family = binomial(link="cloglog"), data = WVS)

model6 <- step(full_model, direction = "both")
rm(full_model)

summary(model6)
```

### 37.

```{r model6 gof test}
pchisq(deviance(model6), df.residual(model6), lower.tail = FALSE)
```

### 38.

```{r model6 sample size check}
model6_df = data_frame()

WVS$model6_preds <- fitted(model6, type="response")
model6_df = data_frame(yes = WVS$yes,
                       no = WVS$no,
                       pred_yes = WVS$total * WVS$model6_preds,
                       pred_no = WVS$total * (1 - WVS$model6_preds))

sum(model6_df$pred_yes < 5)
sum(model6_df$pred_yes < 1)
sum(model6_df$pred_no < 5)
sum(model6_df$pred_no < 1)
```

### 39.

```{r model6 multicolineaity check}
vif(model6)
```

### 40.

```{r model6 residuals check}
residualPlots(model6, type = "deviance", smooth=FALSE)
residualPlots(model6, type = "pearson", smooth=FALSE)
```

### 41.

```{r model6 influential obs check}
p <- length(coef(model6))
n <- nobs(model6)
leverage_cutoff <-  (3*p)/n
cooks_cutoff <- 2/sqrt(n)

model6_inf <- data.frame(
  obs = seq(n),
  studentized = rstudent(model6),
  hat = hatvalues(model6),
  cooks = cooks.distance(model6)
)

model6_inf |>
  ggplot(aes(x = obs, y = studentized)) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.25) +
  geom_hline(yintercept = -3, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_hline(yintercept = 3, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_segment(aes(xend = obs, y = 0, yend = studentized), linewidth = 0.25) +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("Studentized residuals") + 
  theme_bw()

model6_inf |>
  ggplot(aes(x = obs, y = hat)) +
  geom_segment(aes(xend = obs, y = 0, yend = hat), linewidth = 0.25) +
  geom_hline(yintercept = leverage_cutoff, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("hat-values") + 
  theme_bw()

model6_inf |>
  ggplot(aes(x = obs, y = cooks)) +
  geom_segment(aes(xend = obs, y = 0, yend = cooks), linewidth = 0.25) +
  geom_hline(yintercept = cooks_cutoff, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("Cook's distance") + 
  theme_bw()

model6_inf |>
  filter(abs(studentized) > 3 |
         hat > leverage_cutoff |
         cooks > cooks_cutoff
  )
```

### 42.

```{r model6 overdispersion check}
check_overdispersion(model6)
```

## Model 7: loglog, all predictors

### 43.

```{r model7 fit}
model7 <- glm(cbind(no, yes) ~ gender + religion + degree + country + age, family = binomial(link="cloglog"), data = WVS)

summary(model7)
```

### 44.

```{r model7 gof test}
pchisq(deviance(model7), df.residual(model7), lower.tail = FALSE)
```

### 45.

```{r model7 sample size check}
model7_df = data_frame()

WVS$model7_preds <- fitted(model7, type="response")
model7_df = data_frame(yes = WVS$yes,
                       no = WVS$no,
                       pred_yes = WVS$total * WVS$model7_preds,
                       pred_no = WVS$total * (1 - WVS$model7_preds))

sum(model7_df$pred_yes < 5)
sum(model7_df$pred_yes < 1)
sum(model7_df$pred_no < 5)
sum(model7_df$pred_no < 1)
```

### 46.

```{r model7 multicolineaity check}
vif(model7)
```

### 47.

```{r model7 residuals check}
residualPlots(model7, type = "deviance", smooth=FALSE)
residualPlots(model7, type = "pearson", smooth=FALSE)
```

### 48.

```{r model7 influential obs check}
p <- length(coef(model7))
n <- nobs(model7)
leverage_cutoff <-  (3*p)/n
cooks_cutoff <- 2/sqrt(n)

model7_inf <- data.frame(
  obs = seq(n),
  studentized = rstudent(model7),
  hat = hatvalues(model7),
  cooks = cooks.distance(model7)
)

model7_inf |>
  ggplot(aes(x = obs, y = studentized)) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.25) +
  geom_hline(yintercept = -3, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_hline(yintercept = 3, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_segment(aes(xend = obs, y = 0, yend = studentized), linewidth = 0.25) +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("Studentized residuals") + 
  theme_bw()

model7_inf |>
  ggplot(aes(x = obs, y = hat)) +
  geom_segment(aes(xend = obs, y = 0, yend = hat), linewidth = 0.25) +
  geom_hline(yintercept = leverage_cutoff, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("hat-values") + 
  theme_bw()

model7_inf |>
  ggplot(aes(x = obs, y = cooks)) +
  geom_segment(aes(xend = obs, y = 0, yend = cooks), linewidth = 0.25) +
  geom_hline(yintercept = cooks_cutoff, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("Cook's distance") + 
  theme_bw()

model7_inf |>
  filter(abs(studentized) > 3 |
         hat > leverage_cutoff |
         cooks > cooks_cutoff
  )
```

### 49.

```{r model7 overdispersion check}
check_overdispersion(model7)
```

## Model 8: loglog, stepwise selected predictors

### 50.

```{r model8 fit}
full_model <- glm(cbind(no, yes) ~ (gender + religion + degree + country + age)^2, family = binomial(link="cloglog"), data = WVS)

model8 <- step(full_model, direction = "both")
rm(full_model)

summary(model8)
```

### 51.

```{r model8 gof test}
pchisq(deviance(model8), df.residual(model8), lower.tail = FALSE)
```

### 52.

```{r model8 sample size check}
model8_df = data_frame()

WVS$model8_preds <- fitted(model2, type="response")
model8_df = data_frame(yes = WVS$yes,
                       no = WVS$no,
                       pred_yes = WVS$total * WVS$model8_preds,
                       pred_no = WVS$total * (1 - WVS$model8_preds))

sum(model8_df$pred_yes < 5)
sum(model8_df$pred_yes < 1)
sum(model8_df$pred_no < 5)
sum(model8_df$pred_no < 1)
```

### 53.

```{r model8 multicolineaity check}
vif(model8)
```

### 54.

```{r model8 residuals check}
residualPlots(model8, type = "deviance", smooth=FALSE)
residualPlots(model8, type = "pearson", smooth=FALSE)
```

### 55.

```{r model8 influential obs check}
p <- length(coef(model8))
n <- nobs(model8)
leverage_cutoff <-  (3*p)/n
cooks_cutoff <- 2/sqrt(n)

model8_inf <- data.frame(
  obs = seq(n),
  studentized = rstudent(model8),
  hat = hatvalues(model8),
  cooks = cooks.distance(model8)
)

model8_inf |>
  ggplot(aes(x = obs, y = studentized)) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.25) +
  geom_hline(yintercept = -3, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_hline(yintercept = 3, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_segment(aes(xend = obs, y = 0, yend = studentized), linewidth = 0.25) +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("Studentized residuals") + 
  theme_bw()

model8_inf |>
  ggplot(aes(x = obs, y = hat)) +
  geom_segment(aes(xend = obs, y = 0, yend = hat), linewidth = 0.25) +
  geom_hline(yintercept = leverage_cutoff, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("hat-values") + 
  theme_bw()

model8_inf |>
  ggplot(aes(x = obs, y = cooks)) +
  geom_segment(aes(xend = obs, y = 0, yend = cooks), linewidth = 0.25) +
  geom_hline(yintercept = cooks_cutoff, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("Cook's distance") + 
  theme_bw()

model8_inf |>
  filter(abs(studentized) > 3 |
         hat > leverage_cutoff |
         cooks > cooks_cutoff
  )
```

### 56.

```{r model8 overdispersion check}
check_overdispersion(model8)
```

## Model 2, 4, 6, 8 AIC comparisons

### 57.

```{r step aic}
AIC(model2)
AIC(model4)
AIC(model6)
AIC(model8)
```

## Models 9-10: logit, manually reduced variables

### 58.

```{r model2 lrts}
model2_lrt <- Anova(model2)
model2_lrt_p <- model2_lrt[, "Pr(>Chisq)"]

cbind(model2_lrt, p.adj = p.adjust(model2_lrt_p, method = "BH")) |>
  filter(p.adj > 0.05)
```

### 59.

```{r model9 fit}
model9 <- update(model2, . ~ . - gender:degree)
summary(model9)
```

### 60.

```{r model9 vs model2 lrt}
anova(model2, model9)
```

### 61.

```{r model9 lrts}
model9_lrt <- Anova(model9)
model9_lrt_p <- model9_lrt[, "Pr(>Chisq)"]

cbind(model9_lrt, p.adj = p.adjust(model9_lrt_p, method = "BH")) |>
  filter(p.adj > 0.05)
```

### 62.

```{r model10 fit}
model10 <- update(model9, . ~ . - religion:age)
summary(model10)
```

### 63.

```{r model10 vs model9 lrt}
anova(model9, model10)
```

### 64.

```{r model10 lrts}
model10_lrt <- Anova(model10)
model9_lrt_10 <- model10_lrt[, "Pr(>Chisq)"]

cbind(model10_lrt, p.adj = p.adjust(model9_lrt_10, method = "BH")) |>
  filter(p.adj > 0.05)
```

### 65.

```{r model10 gof test}
pchisq(deviance(model10), df.residual(model10), lower.tail = FALSE)
```

### 66.

```{r model10 sample size check}
model10_df = data_frame()

WVS$model10_preds <- fitted(model10, type="response")
model10_df = data_frame(yes = WVS$yes,
                       no = WVS$no,
                       pred_yes = WVS$total * WVS$model10_preds,
                       pred_no = WVS$total * (1 - WVS$model10_preds))

sum(model10_df$pred_yes < 5)
sum(model10_df$pred_yes < 1)
sum(model10_df$pred_no < 5)
sum(model10_df$pred_no < 1)
```

### 67.

```{r model10 multicolineaity check}
vif(model10)
```

### 68.

```{r model10 residuals check}
residualPlots(model10, type = "deviance", smooth=FALSE)
residualPlots(model10, type = "pearson", smooth=FALSE)
```

### 69.

```{r model10 influential obs check}
p <- length(coef(model10))
n <- nobs(model10)
leverage_cutoff <-  (3*p)/n
cooks_cutoff <- 2/sqrt(n)

model10_inf <- data.frame(
  obs = seq(n),
  studentized = rstudent(model10),
  hat = hatvalues(model10),
  cooks = cooks.distance(model10)
)

model10_inf |>
  ggplot(aes(x = obs, y = studentized)) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.25) +
  geom_hline(yintercept = -3, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_hline(yintercept = 3, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_segment(aes(xend = obs, y = 0, yend = studentized), linewidth = 0.25) +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("Studentized residuals") + 
  theme_bw()

model10_inf |>
  ggplot(aes(x = obs, y = hat)) +
  geom_segment(aes(xend = obs, y = 0, yend = hat), linewidth = 0.25) +
  geom_hline(yintercept = leverage_cutoff, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("hat-values") + 
  theme_bw()

model10_inf |>
  ggplot(aes(x = obs, y = cooks)) +
  geom_segment(aes(xend = obs, y = 0, yend = cooks), linewidth = 0.25) +
  geom_hline(yintercept = cooks_cutoff, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("Cook's distance") + 
  theme_bw()

model10_inf |>
  filter(abs(studentized) > 3 |
         hat > leverage_cutoff |
         cooks > cooks_cutoff
  )
```

### 70.

```{r model10 overdispersion check}
check_overdispersion(model10)
```

## Models 11-12: probit, manually reduced variables

### 71.

```{r model4 lrts}
model4_lrt <- Anova(model4)
model4_lrt_p <- model4_lrt[, "Pr(>Chisq)"]

cbind(model4_lrt, p.adj = p.adjust(model4_lrt_p, method = "BH")) |>
  filter(p.adj > 0.05)
```

### 72.

```{r model11 fit}
model11 <- update(model4, . ~ . - gender:degree)
summary(model11)
```

### 73.

```{r model11 vs model4 lrt}
anova(model4, model11)
```

### 74.

```{r model11 lrts}
model11_lrt <- Anova(model11)
model11_lrt_p <- model11_lrt[, "Pr(>Chisq)"]

cbind(model11_lrt, p.adj = p.adjust(model11_lrt_p, method = "BH")) |>
  filter(p.adj > 0.05)
```

### 75.

```{r model12 fit}
model12 <- update(model11, . ~ . - religion:age)
summary(model12)
```

### 76.

```{r model12 vs model11 lrt}
anova(model11, model12)
```

### 77.

```{r model12 lrts}
model12_lrt <- Anova(model12)
model12_lrt_p <- model12_lrt[, "Pr(>Chisq)"]

cbind(model12_lrt, p.adj = p.adjust(model12_lrt_p, method = "BH")) |>
  filter(p.adj > 0.05)
```

### 78.

```{r model12 gof test}
pchisq(deviance(model12), df.residual(model12), lower.tail = FALSE)
```

### 79.

```{r model12 sample size check}
model12_df = data_frame()

WVS$model12_preds <- fitted(model12, type="response")
model12_df = data_frame(yes = WVS$yes,
                       no = WVS$no,
                       pred_yes = WVS$total * WVS$model12_preds,
                       pred_no = WVS$total * (1 - WVS$model12_preds))

sum(model12_df$pred_yes < 5)
sum(model12_df$pred_yes < 1)
sum(model12_df$pred_no < 5)
sum(model12_df$pred_no < 1)
```

### 80.

```{r model12 multicolineaity check}
vif(model12)
```

### 81.

```{r model12 residuals check}
residualPlots(model12, type = "deviance", smooth=FALSE)
residualPlots(model12, type = "pearson", smooth=FALSE)
```

### 82.

```{r model12 influential obs check}
p <- length(coef(model12))
n <- nobs(model12)
leverage_cutoff <-  (3*p)/n
cooks_cutoff <- 2/sqrt(n)

model12_inf <- data.frame(
  obs = seq(n),
  studentized = rstudent(model12),
  hat = hatvalues(model12),
  cooks = cooks.distance(model12)
)

model12_inf |>
  ggplot(aes(x = obs, y = studentized)) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.25) +
  geom_hline(yintercept = -3, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_hline(yintercept = 3, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_segment(aes(xend = obs, y = 0, yend = studentized), linewidth = 0.25) +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("Studentized residuals") + 
  theme_bw()

model12_inf |>
  ggplot(aes(x = obs, y = hat)) +
  geom_segment(aes(xend = obs, y = 0, yend = hat), linewidth = 0.25) +
  geom_hline(yintercept = leverage_cutoff, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("hat-values") + 
  theme_bw()

model12_inf |>
  ggplot(aes(x = obs, y = cooks)) +
  geom_segment(aes(xend = obs, y = 0, yend = cooks), linewidth = 0.25) +
  geom_hline(yintercept = cooks_cutoff, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("Cook's distance") + 
  theme_bw()

model12_inf |>
  filter(abs(studentized) > 3 |
         hat > leverage_cutoff |
         cooks > cooks_cutoff
  )
```

### 83

```{r model12 overdispersion check}
check_overdispersion(model12)
```

## Models 13-14: cloglog, manually reduced variables

### 84.

```{r model6 lrts}
model6_lrt <- Anova(model6)
model6_lrt_p <- model6_lrt[, "Pr(>Chisq)"]

cbind(model6_lrt, p.adj = p.adjust(model6_lrt_p, method = "BH")) |>
  filter(p.adj > 0.05)
```

### 85.

```{r model13 fit}
model13 <- update(model6, . ~ . - gender:degree)
summary(model13)
```

### 86.

```{r model13 vs model6 lrt}
anova(model6, model13)
```

### 87.

```{r model13 lrts}
model13_lrt <- Anova(model13)
model13_lrt_p <- model13_lrt[, "Pr(>Chisq)"]

cbind(model13_lrt, p.adj = p.adjust(model13_lrt_p, method = "BH")) |>
  filter(p.adj > 0.05)
```

### 88.

```{r model14 fit}
model14 <- update(model13, . ~ . - religion:age)
summary(model14)
```

### 89.

```{r model14 vs model13 lrt}
anova(model13, model14)
```

### 90.

```{r model14 lrts}
model14_lrt <- Anova(model14)
model14_lrt_p <- model14_lrt[, "Pr(>Chisq)"]

cbind(model14_lrt, p.adj = p.adjust(model14_lrt_p, method = "BH")) |>
  filter(p.adj > 0.05)
```

### 91.

```{r model14 gof test}
pchisq(deviance(model14), df.residual(model14), lower.tail = FALSE)
```

### 92.

```{r model14 sample size check}
model14_df = data_frame()

WVS$model14_preds <- fitted(model14, type="response")
model14_df = data_frame(yes = WVS$yes,
                       no = WVS$no,
                       pred_yes = WVS$total * WVS$model14_preds,
                       pred_no = WVS$total * (1 - WVS$model14_preds))

sum(model14_df$pred_yes < 5)
sum(model14_df$pred_yes < 1)
sum(model14_df$pred_no < 5)
sum(model14_df$pred_no < 1)
```

### 93.

```{r model14 multicolineaity check}
vif(model14)
```

### 94.

```{r model14 residuals check}
residualPlots(model14, type = "deviance", smooth=FALSE)
residualPlots(model14, type = "pearson", smooth=FALSE)
```

### 95.

```{r model14 influential obs check}
p <- length(coef(model14))
n <- nobs(model14)
leverage_cutoff <-  (3*p)/n
cooks_cutoff <- 2/sqrt(n)

model14_inf <- data.frame(
  obs = seq(n),
  studentized = rstudent(model14),
  hat = hatvalues(model14),
  cooks = cooks.distance(model14)
)

model14_inf |>
  ggplot(aes(x = obs, y = studentized)) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.25) +
  geom_hline(yintercept = -3, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_hline(yintercept = 3, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_segment(aes(xend = obs, y = 0, yend = studentized), linewidth = 0.25) +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("Studentized residuals") + 
  theme_bw()

model14_inf |>
  ggplot(aes(x = obs, y = hat)) +
  geom_segment(aes(xend = obs, y = 0, yend = hat), linewidth = 0.25) +
  geom_hline(yintercept = leverage_cutoff, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("hat-values") + 
  theme_bw()

model14_inf |>
  ggplot(aes(x = obs, y = cooks)) +
  geom_segment(aes(xend = obs, y = 0, yend = cooks), linewidth = 0.25) +
  geom_hline(yintercept = cooks_cutoff, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("Cook's distance") + 
  theme_bw()

model14_inf |>
  filter(abs(studentized) > 3 |
         hat > leverage_cutoff |
         cooks > cooks_cutoff
  )
```

### 96.

```{r model14 overdispersion check}
check_overdispersion(model14)
```

## Models 15-16: loglog, manually reduced variables

### 97.

```{r model8 lrts}
model8_lrt <- Anova(model8)
model8_lrt_p <- model8_lrt[, "Pr(>Chisq)"]

cbind(model8_lrt, p.adj = p.adjust(model8_lrt_p, method = "BH")) |>
  filter(p.adj > 0.05)
```

### 98.

```{r model15 fit}
model15 <- update(model8, . ~ . - gender:degree)
summary(model15)
```

### 99.

```{r model15 vs model8 lrt}
anova(model8, model15)
```

### 100.

```{r model15 lrts}
model15_lrt <- Anova(model15)
model15_lrt_p <- model15_lrt[, "Pr(>Chisq)"]

cbind(model15_lrt, p.adj = p.adjust(model15_lrt_p, method = "BH")) |>
  filter(p.adj > 0.05)
```

### 101.

```{r fit model16}
model16 <- update(model15, . ~ . - religion:age)
summary(model16)
```

### 102.

```{r model16 vs model15 lrt}
anova(model15, model16)
```

### 103.

```{r model16 lrts}
model16_lrt <- Anova(model16)
model16_lrt_p <- model16_lrt[, "Pr(>Chisq)"]

cbind(model16_lrt, p.adj = p.adjust(model16_lrt_p, method = "BH")) |>
  filter(p.adj > 0.05)
```

### 104.

```{r model16 gof test}
pchisq(deviance(model16), df.residual(model16), lower.tail = FALSE)
```

### 105.

```{r model16 sample size check}
model16_df = data_frame()

WVS$model16_preds <- fitted(model16, type="response")
model16_df = data_frame(yes = WVS$yes,
                       no = WVS$no,
                       pred_yes = WVS$total * WVS$model16_preds,
                       pred_no = WVS$total * (1 - WVS$model16_preds))

sum(model16_df$pred_yes < 5)
sum(model16_df$pred_yes < 1)
sum(model16_df$pred_no < 5)
sum(model16_df$pred_no < 1)
```

### 106.

```{r model16 multicolineaity check}
vif(model16)
```

### 107.

```{r model16 residuals check}
residualPlots(model16, type = "deviance", smooth=FALSE)
residualPlots(model16, type = "pearson", smooth=FALSE)
```

### 108.

```{r model16 influential obs check}
p <- length(coef(model16))
n <- nobs(model16)
leverage_cutoff <-  (3*p)/n
cooks_cutoff <- 2/sqrt(n)

model16_inf <- data.frame(
  obs = seq(n),
  studentized = rstudent(model16),
  hat = hatvalues(model16),
  cooks = cooks.distance(model16)
)

model16_inf |>
  ggplot(aes(x = obs, y = studentized)) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.25) +
  geom_hline(yintercept = -3, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_hline(yintercept = 3, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_segment(aes(xend = obs, y = 0, yend = studentized), linewidth = 0.25) +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("Studentized residuals") + 
  theme_bw()

model16_inf |>
  ggplot(aes(x = obs, y = hat)) +
  geom_segment(aes(xend = obs, y = 0, yend = hat), linewidth = 0.25) +
  geom_hline(yintercept = leverage_cutoff, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("hat-values") + 
  theme_bw()

model16_inf |>
  ggplot(aes(x = obs, y = cooks)) +
  geom_segment(aes(xend = obs, y = 0, yend = cooks), linewidth = 0.25) +
  geom_hline(yintercept = cooks_cutoff, linetype = "dashed", linewidth = 0.25, color = "red") +
  geom_point(shape = 1) +
  xlab("Index") +
  ylab("Cook's distance") + 
  theme_bw()

model16_inf |>
  filter(abs(studentized) > 3 |
         hat > leverage_cutoff |
         cooks > cooks_cutoff
  )
```

### 109.

```{r model16 overdispersion check}
check_overdispersion(model16)
```

## Model 10, 12, 14, 16 AIC comparisons

### 110.

```{r final aic}
AIC(model10)
AIC(model12)
AIC(model14)
AIC(model16)
```

## Final Model: Model 10

### 111.

```{r model10 vs full model lrt}
full_model <- glm(cbind(yes, no) ~ (gender + religion + degree + country + age)^2, family = binomial(link="logit"), data = WVS)

anova(model10, full_model)
```

```{r final model}
summary(model10)
```

### 112.

```{r final model predictions}
WVS$final_fits <- fitted(model10)
```

### 113.

```{r final model roc-auc}
model10_ungrp <- glm(formula = resp ~ gender + religion + degree + country + age + religion:degree + religion:country + degree:country + country:age, family = binomial(link = "logit"), data = WVS_ungrouped)


WVS_ungrouped$final_fits <- fitted(model10_ungrp)

roc <- roc(data = WVS_ungrouped,
                  response = resp,
                  predictor = final_fits,
                  auc=TRUE, plot=TRUE)
auc(roc)
```

### 114.

```{r final model predictive power}
WVS_ungrouped$pred_resp <- ifelse(WVS_ungrouped$final_fits > 0.5, 1, 0)

WVS_ungrouped$pred_resp <- factor(WVS_ungrouped$pred_resp, levels = c(0, 1))
WVS_ungrouped$resp <- factor(WVS_ungrouped$resp, levels = c(0, 1))

confusionMatrix(data = WVS_ungrouped$pred_resp, reference = WVS_ungrouped$resp, positive = "1")
```
